---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl) #read excel files
library(haven) #read SAS files
```

## Read in some data

Read in the litters dataset.

```{r}
litters_df = read_csv("./data/FAS_litters.csv") #complex names bother us
litters_df = janitor::clean_names(litters_df) #can use library(janitor), but at this time, we only need one function (clean_names) which from the janitor package, so we can use janitor::clean_names to import the certain function
```

## Take a look at the data

Printing in the console.

```{r}
litters_df
```

```{r}
head(litters_df) #first 6 rows

tail(litters_df) #last 6 rows
```

```{r}
skimr::skim(litters_df)
```

## Options to read_csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE) #skip first 10 rows
litters_df
```

```{r}
litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", ".", 999))
litters_df
```

```{r}
#litters_df = read_csv("./data/FAS_litters.csv", col_types = )
#litters_df
```

check out `?read_csv()` for more information.

## Other file formats

Read in an excel file.

```{r}
library(readxl) #read excel files
```

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

Read in a SAS file.

```{r}
library(haven) #read SAS files
```

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with Base R

what about `read.csv`...? 
read.csv is build into R, don't need to load any packages

use read_csv! do NOT use read.csv!!!

```{r}
litters_base = read.csv("data/FAS_litters.csv")
litters_readr = read_csv("data/FAS_litters.csv")

litters_base #more difficult to read
litters_readr
```

## Exporting data

Exporting the mlb sub-table.

```{r}
write_csv(mlb_df, "./data/mlb_subtable.csv")
```




